---
title: "Clustering Homework"
output:
  html_document:
    df_print: paged
---


```{r}
library(tidyverse)
library(janitor)
library(broom)
library(factoextra)
```

<br>

--------------------------------------------------------------------------------

<br>

# MVP

<br>

## 1. Explore Data 

<br>



```{r}
mall_customers <- read_csv("data/mall_customers.csv") %>% clean_names()
```

```{r}
mall_customers 
```
there are 2 non-numerical columns: customerID and gender
k means method uses numerical data
remove CustomerID column and gender column will most likely be irrelevant 

```{r}
mall_customers_numerical <- mall_customers %>% 
  select(-customer_id, -gender)
```

Visualise variables in question (annual income and spending score)
```{r}
mall_customers_numerical %>% 
  ggplot(aes(x = annual_income_k, y = spending_score_1_100)) +
  geom_point()
```

It does look like there are some potential groups so this data would be appropriate for clustering
You can see one very 'tight' cluster around 50,50, surrounded by 4 'looser' clusters 


<br>

## 2.

### Scaling 

First, check what we currently have for mean and SD in the two variables

```{r}
mall_customers_numerical %>% 
  as_tibble() %>% 
  pivot_longer(cols = c("annual_income_k", "spending_score_1_100"),
               names_to = "data_category",
               values_to = "value") %>% 
  group_by(data_category) %>% 
  summarise(mean = round(mean(value)),
            sd = sd(value))
  
```
We need the data to be scaled so that the mean is 0 and the sd is one

```{r}
(mall_scale <- mall_customers_numerical %>% 
   select(annual_income_k, spending_score_1_100) %>% 
    mutate(across(everything(),
         .fns = scale))
)
```

Check scaled mean and sd are 0 and 1:

```{r}
mall_scale %>% 
  as_tibble() %>% 
  pivot_longer(cols = c("annual_income_k", "spending_score_1_100"),
               names_to = "data_category",
               values_to = "value") %>% 
  group_by(data_category) %>% 
  summarise(mean = round(mean(value)),
            sd = sd(value))
```

<br>


### Clustering

```{r}
(clustered_mall <- kmeans(x = mall_scale,
                         centers = 5,
                         nstart = 25)
)
```

```{r}
min_k <- 1
max_k <- 20

k_clusters <- tibble(k = min_k:max_k) %>% 
  mutate(
    kclust = map(k, ~kmeans(mall_scale, .x, nstart = 25)),
    tidied = map(kclust, tidy),
    glanced = map(kclust, glance),
    augmented = map(kclust, augment, mall_customers_numerical)
  )

k_clusters
```

```{r}
(clusterings <- k_clusters %>% 
  unnest(glanced)
)
```

```{r}
clusterings %>% 
  ggplot(aes(x = k, y = tot.withinss)) +
  geom_point() +
  geom_line() +
  scale_x_continuous(breaks = seq(1, 20, by = 1))
```

There is a kink at 3 and at 5, check this with fviz 'silhouette'

```{r}
fviz_nbclust(mall_scale,
             kmeans,
             method = "silhouette",
             nstart = 25)
```

A value of k = 5 is the best fit

<br>


## 3. Visualise clustering with k = 5

```{r}
clusterings %>% 
  unnest(cols = c(augmented)) %>% 
  filter(k == 5) %>% 
  ggplot(aes(x = annual_income_k, y = spending_score_1_100)) +
  geom_point(aes(colour = .cluster)) 
  
```


<br>


## 4. Does the clustering seem like a good fit for this data?

Check mean for each cluster

```{r}
clusterings %>% 
  unnest(augmented) %>% 
  filter(k == 5) %>% 
  group_by(.cluster) %>% 
  summarise(mean(annual_income_k), mean(spending_score_1_100))
```

Some of the means for annual_income are quite similar (clusters 1 & 4 and clusters 2 & 5) as well as some clusters (1 & 5 and 2 & 4) from the spending_score, so perhaps they could have been clustered together better? However visually the data looks appropriate for clustering as you can clearly see definied clusters on the plot.

<br>


## 5. Cluster Attributes


Assuming that a high spending score indicates that the customer spends a lot of money in the mall:

- Clusters 1 and 4 have a low annual income but a high spending score ("super fans")
- Clusters 2 and 5 have a high annual income but a low spending score ("tag alongs")
- Cluster 5 has a mid-range income and spending score ("soloists")

