---
title: "W2D2_Homework_Kirsty"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---

--------------------------------------------------------------------------------

# Load Libraries & Data

```{r}
library(tidyverse)

```

```{r}
info   <- read_csv("data/code_clan_info.csv")
tweets <- read_csv("data/code_clan_tweets.csv")
```


--------------------------------------------------------------------------------


# MVP

## Question 1:

#### Load the code_clan_tweets.csv data. Find the number of rows, columns, and list all the variable names.

```{r}
dim(tweets)
```

-   There are 234 rows and 27 columns.
-   The variable names are:

```{r}
colnames(tweets)
```

## Question 2:

#### Find the total number of favourites (stored in favorite_count) that CodeClan tweets have got. Don't add a tweet's favorite_count to the total when the tweet was quoted (see the is_quote variable).

```{r}
tweets_minus_quotes <-
  tweets %>% 
  filter(is_quote == "FALSE")

tweets_minus_quotes %>% 
  summarise(favorite_count_not_quoted = sum(favorite_count))


```

-   The total number of favourites that CodeClan tweets got was 425.

## Question 3:

#### Summarise the mean number of retweets (retweet_count), split by the type of platform that was used to tweet (source). Again, omit quoted tweets from the mean.

```{r}
# check sources
tweets_minus_quotes %>% 
  count(source)


```

```{r}
tweets_minus_quotes %>% 
  group_by(source) %>% 
  summarise(mean_retweets = mean(retweet_count)) 


```

## Question 4:

#### Count the total number of likes (i.e. total of favorite_count), split by media type, and arrange them from most likes to least. Assume that any tweets without a listed media type are of type "text".

```{r}
# check media types
tweets %>% 
  count(media_type)
```

```{r}
tweets %>% 
  mutate(media_imputed = coalesce(media_type, "text"),
         .after = media_type) %>% 
  group_by(media_imputed) %>% 
  summarise(total_favorite_count = sum(favorite_count)) %>% 
  arrange(desc(total_favorite_count))
  
```

## Question 5:

#### Find the mean number of characters that a CodeClan tweet contains. You can either find a way to count the text in the text variable, or use the variable display_text_width which contains a count. Checking both would be a good idea if possible to see if they match (but it's not essential).

```{r}
tweets %>% 
  mutate(tweet_characters = nchar(text),
         .after = display_text_width) %>%  
  # the display_text_width column and the tweet_characters column do not match.
  # I will use the tweet_characters column to find the mean as I know it's accurate!
  summarise(mean_characters = mean(tweet_characters))
```

-   The mean number of characters that a CodeClan tweet contains is 188.

## Question 6:

#### The code_clan_info.csv data file contains status_url along with other info. Load this in, and join it to the code_clan_tweets tibble, so that you have a status_url for each tweet. Decide which variable to join the tibbles on.

```{r}
head(info)

#both tables contain a "tweet_id" column, so I will use this to join
```

```{r}
tweets_info <- tweets %>% 
  left_join(info, "tweet_id") %>% 
  select(tweet_id:retweet_location, status_url)

tweets_info
```

## Question 7:

#### From your new joined data, create a new tibble codeclan_hashtags containing only tweet_id and hashtags, then convert hashtags into lowercase for analysis. Keep only those tweets with hashtags.

```{r}
codeclan_hashtags <- tweets_info %>% 
  select(tweet_id, hashtags) %>% 
  mutate(hashtags = (str_to_lower(hashtags))) %>% 
  drop_na(hashtags)

codeclan_hashtags
```


--------------------------------------------------------------------------------

# Extension 

## Question 8:

#### Some tweets have more than one hashtag, denoted by the c( symbols at the start of the string. Use the str_sub() and str_detect() functions and your codeclan_hashtags tibble from above to find all the cases in which the hashtag string begins with characters c(.

```{r}
codeclan_hashtags %>% 
 mutate(multiple_hashtages = str_sub(hashtags,
          start = 1,
          end   = 2)) %>% 
  mutate(multiple_hashtages = str_detect(multiple_hashtages, "c\\(")) %>% 
  filter(multiple_hashtages == "TRUE")
```

## Question 9:
### Use the str_detect() function to find all tweets with text that mentions “Edinburgh”, and count how many such tweets there are.
 
```{r}
tweets %>% 
  mutate(edinburgh = str_detect(text, "[eE]dinburgh"),
         .after = text) %>% 
  count(edinburgh)
```
- There are 33 tweets that mention "Edinburgh"

# Question 10: 

#### Use str_extract_all() with a regular expression to find out which Twitter users CodeClan have been tweeting.

```{r}
tweets %>% 
  mutate(tweets_to = str_extract_all(reply_to_screen_name, "[A-za-z0-9_]*"),
         .after = reply_to_screen_name) %>% 
  unnest(tweets_to)
```

