---
title: "wknd2_meteorites_kirsty"
output: html_notebook
---
--------------------------------------------------------------------------------

# Load libraries & data 

```{r}
library(tidyverse)
library(testthat)
```

```{r}
meteorites_cleaned_filtered <- read_csv("data/meteorites_cleaned_filtered.csv")
```

# 1.2 The .Rmd file

## Q2: Find the names and years found for the 10 largest meteorites in the data.
```{r}
meteorites_cleaned_filtered %>% 
  select(name, mass_g, year) %>% #select columns required
  arrange(desc(mass_g)) %>% 
  head(10)
```





## Q3: Find the average mass of meteorites that were recorded falling, vs. those which were just found.
```{r}
# check how many types of "fall" there are
meteorites_cleaned_filtered %>% 
  count(fall)
```
```{r}
meteorites_cleaned_filtered %>% 
  group_by(fall) %>% 
  summarise(average_weight = mean(mass_g))
```





## Q4: Find the number of meteorites in each year, for every year since 2000.

```{r}
meteorites_cleaned_filtered %>% 
  filter(year >= 2000) %>% 
  group_by(year) %>%
  summarise(total_meteorites = n()) 
```





## Q5: Write a function called get_meteorite_count() that takes in two arguments: a dataframe and a year, and returns the phrase “In (input_year), there were (number of meteorites that year) meteorites.” IF there are 1 or more meteorites that year. If there are no meteorites in the year, return the phrase “As far as we know, there were no meteorites that year”.

```{r}
# filter dataframe for input year, then count total of meteorites in that year
get_meteorite_count <- function(dataframe, input_year){
  total_meteorites <- dataframe %>% 
    filter(year == input_year) %>%
    summarise(total_meteorites = n()) %>% 
    pull(total_meteorites)
  
  # If total = 0, then state so; otherwise state total
  if (total_meteorites == 0) {
    print("As far as we know, there were no meteorites that year")
  } else { if (total_meteorites >= 1){
    print(paste0("In ", input_year, ", there were ", total_meteorites, " meteorites."))
  }
  }
}

  
```

```{r}
# test new function

get_meteorite_count(meteorites_cleaned_filtered, 2000)

```





## Q6: Write down ideas for how the function you’ve written could be improved and how you could test that the function works as it should. 

- you could test that the dataframe does have the variables "year" and "total_meteorites"
- you could check that both of these variables are integers
- you could add a failure statement for non-numeric input
- you could add a test for logical input
- you could check that the year inputted does exist in the dataframe
- you could provide some extra information, such as the largest meteorite in the year


## Q7: Write a unit test for your function to check whether non-numeric inputs (character and logical) for the year argument return an error.


```{r}
# using the `test_that` package

test_that("Non-numeric year inputs returns an error", {
  expect_error(get_meteorite_count(meteorites_cleaned_filtered, year))
  expect_error(get_meteorite_count(meteorites_cleaned_filtered, TRUE))
})

```
  - Since `TRUE` is logical, it is seen as binary and therefore represents a number, so can pass through the function.





## Q8: Improve your function based on the result of your test. Instead of going back and changing the function, please create a new function that is the improved version called refactor_get_meteorite_count(). Test your function to make sure it now gives an error when you input a logical or character value as the year. 

```{r}

# filter dataframe for input year, then count total of meteorites in that year
refactor_get_meteorite_count <- function(dataframe, input_year){
  total_meteorites <- dataframe %>% 
    filter(year == input_year) %>%
    summarise(total_meteorites = n()) %>% 
    pull(total_meteorites)
  
# Check if input_year is numerical, stop if not
  
  if (!is.numeric(input_year)) {
    stop("Non-numeric input")
  } 
  
# If total = 0, then state so; otherwise state total
  
  if (total_meteorites == 0) {
    print("As far as we know, there were no meteorites that year")
  } else { if (total_meteorites >= 1){
    print(paste0("In ", input_year, ", there were ", total_meteorites, " meteorites."))
  }
  }
}
```

```{r}
# Test new function
refactor_get_meteorite_count(meteorites_cleaned_filtered, TRUE)
```


