---
title: "Model Building Weekend Homework"
output: html_notebook
---


```{r}
library(tidyverse)
library(GGally)
library(ggfortify)
library(modelr)
```

```{r}
red_wine <- read_csv("data/wine_quality_red.csv")

white_wine <- read_csv("data/wine_quality_white.csv")

head(red_wine)
head(white_wine)
```

```{r}
# check for NA values 
red_wine %>% 
  filter(across(everything(), ~is.na(.)))

white_wine %>% 
  filter(across(everything(), ~is.na(.)))
```

```{r}
# change wine_id to be id_r for red wine and id_w for white wine
red_wine_clean <- red_wine %>% 
  mutate(wine_id = paste0(wine_id,"_red"))

white_wine_clean <- white_wine %>% 
  mutate(wine_id = paste0(wine_id,"_white"))
```

```{r}
# check col names are the same
colnames(white_wine_clean)
colnames(red_wine_clean)
```

```{r}
# join data sets

(wine_full <- rbind(red_wine_clean, white_wine_clean))
```
```{r}
wine_full %>% 
  ggplot(aes(x = quality)) +
  geom_histogram()
```
data set looks fairly well distributed


data dictionary suggests that good predictors could be:
volatile acidity (high = unpleasant)
citric acid (high = fresher)
residual sugar (high = sweet)
chlorides (salt, assume high = bad?)
total_sulfur_dioxide (low = oxidation and microbial growth (bad?))

at this point, could possibly also drop:
density
pH
sulphates
alcohol
region
as data dictionary does not provide any information on these variables having a distinct impact on taste, but will investigate data set first before committing to dropping these

to drop at this point:
total_sulfur_dioxide is calculated from free_sulfur_dioxide, so it makes sense to drop free_sulfur_dioxide
fixed_acidity seems like it would be related to pH, so drop that
id column

```{r}
(wine_full_clean <- wine_full %>% 
   select(-c("free_sulfur_dioxide", "fixed_acidity", "wine_id")))
```


citric acid column seems very low numbers from 0 - 1.66 
group these into bins of:

< 0.5
0.5 - 1
1 - 1.5
> 1.5

```{r}
(wine_full_clean <- wine_full_clean %>%  
  mutate(citric_acid = case_when(citric_acid < 0.50 ~ "<0.50",
                                    between(citric_acid, 0.50, 0.99) ~ "0.50 - 0.99",
                                    between(citric_acid, 1.00, 1.50) ~ "1.00 - 1.50",
                                    citric_acid > 1.50 ~ ">1.50"), .after = citric_acid
  )) 

```



alias check 

```{r}
wine_sample <- wine_full_clean %>% 
  slice_sample(prop = 0.1) 
  alias(lm(quality ~ ., data = wine_sample))
```
Looks like no predictors are colinear!

```{r message=FALSE, warning=FALSE}
wine_sample %>% 
  ggpairs(progress = FALSE)
```
- It looks like the highest predictor of quality is alcohol content
- other predictors could be volatile acids, possibly citric acid?
- ph and sulphates have a lower correlation but could still be potential predictors
- region does not look like it provides much of an effect
- there is a high correlation between:
  - density and residual salt, which makes sense (why then do these not appear as colinear?)
  - chrlorides and volatile acids
  - alcohol and residual sugars
  - alcohol and density

Start with a SLR between quality and alcohol

```{r}
model_1a <- lm(quality ~ alcohol, data = wine_full_clean)
model_1b <- lm(quality ~ volatile_acidity, data = wine_full_clean)
model_1c <- lm(quality ~ citric_acid, data = wine_full_clean)
```

```{r}
summary(model_1a)
summary(model_1b)
summary(model_1c)
```

```{r}
autoplot(model_1a)
autoplot(model_1b)
autoplot(model_1c)
```
model 1a:
Residuals vs fitted: appears to be randomly scattered around 0
Normal QQ: relatively flat along line with some distance at tails
Scale-Location: no evidence of heteroskedacitity

model 1b:
positively skewed
check if improves when using log(volatile acidity)

model1c:
all plots seem fine

```{r}
model_1b_log <- lm(quality ~ log(volatile_acidity), data = wine_full_clean)
summary(model_1b_log)
autoplot(model_1b_log)

```
This does appear to fix the problem


quality ~ alcohol (model_1a) does appear to have the highest r squared value, so this will be the first predictor

Second predictor:

```{r}
(wine_resid <- wine_full_clean %>% 
  add_residuals(model = model_1a, var = "resid") %>% 
  select(-quality, -alcohol))
```

```{r message=FALSE, warning=FALSE}
wine_resid %>% 
  ggpairs(progress = FALSE)
```
Residual sugars and total_sulfur_dioxide, and sulphates have a correlation with the residuals, investigate adding these
residuals are nicely distributed


```{r}
model_2a <- lm(quality ~ alcohol + residual_sugar, data = wine_full_clean)
model_2b <- lm(quality ~ alcohol + total_sulfur_dioxide, data = wine_full_clean)
model_2c <- lm(quality ~ alcohol + sulphates, data = wine_full_clean)
```


```{r}
summary(model_2a)
summary(model_2b)
summary(model_2c)
```

```{r}
autoplot(model_2a)
autoplot(model_2b)
autoplot(model_2c)
```
- all graphs appear ok
- residual sugar gave the highest r.sq value (model_2a), and improved the r.sq of the SLR(0.1832) so will use this as second predictor

does adding a third predictor help?

```{r message=FALSE, warning=FALSE}
wine_resid_2 <- wine_full_clean %>% 
  add_residuals(model_2a, var = "resid") %>% 
  select(-quality, - alcohol, -residual_sugar)

wine_resid_2 %>% 
  ggpairs(progress = FALSE)
```

- again, residuals look nicely correlated
- sulphates has the highest correlation, investigate adding this to model

```{r}
model_3a <- lm(quality ~ alcohol + residual_sugar + sulphates, data = wine_full_clean)

summary(model_3a)
autoplot(model_3a)
```
plots look ok
r squared value has improved from before (0.1986 in previous model), RSE very slightly reduced, p Value is low for all predictors.
Therefore would say that the best model for prediction is model_3a:

quality ~ alcohol + residual_sugar + sulphates

y = 1.582 + (alcohol * 0.372) + (residual_sugar * 0.028) + (sulphates * 0375)