---
title: "Rates of Cancer Incidences in NHS Borders"
output:
  html_document:
    df_print: paged
---


```{r setup, include=FALSE}
 
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE) 

```

```{r}
# load packages
library(tidyverse)
library(janitor)
library(leaflet)
library(here)

# load data (7 datasets)
geography_codes <- read_csv(here("data/geography_codes_and_labels_hb2014_01042019.csv")) %>%  clean_names() # HB code and HB name
cancer_network_region_annual_incidence <- read_csv(here("data/annual_cancer_incidence_cancer_network_region.csv"))  %>%  clean_names() # region, cancer type, sex, age counts, crude 
scotland_annual_incidence <- read_csv(here("data/annual_cancer_incidence_scotland.csv"))  %>%  clean_names()
healthboard_incidence <- read_csv(here("data/incidence_by_hb.csv"))  %>%  clean_names()
cancer_network_region_5yr_incidence <- read_csv(here("data/five_yr_summary_incidence_cancer_network_region.csv"))  %>%  clean_names()
hb_5yr_incidence <- read_csv(here("data/five_yr_summary_incidence_hb.csv"))  %>%  clean_names()
scotland_5yr_incidence <- read_csv(here("data/five_yr_summary_incidence_scotland.csv"))  %>%  clean_names()

```



<br>

## Introduction

<br>

- This report aims to explore cancer rates in the NHS Borders healthboard.
- Data was gathered from Public Health Scotland, and looks at the cancer rate for numerous cancer types across ages from <5 to 85+ from 2017 to 2021
- The cancer "rate" is the average annual number of cases per population 


<br>



-------------------------------------------------------------------------------

<br>

### Data Quality and Bias

<br>
 
The data is completely anonymised and does not provide any connection to sensitive patient data.

<br>

------------------------------------------------------------------------

<br>

## Data Visualisations

<br>

### **What are the most reported cancers in NHS Borders?**

<br>



```{r}
hb_5yr_incidence %>% 
  filter(hb == "S08000016",
         sex != "All") %>% 
  group_by(cancer_site, sex) %>% 
  summarise(count = sum(incidences_all_ages)) %>% 
  arrange(desc(count)) %>% 
  filter(cancer_site != "All cancer types") %>% 
  head(10) %>%
  mutate(sex = ifelse(sex == "Females", "Female", sex)) %>% 
  ggplot(aes(x = cancer_site, y = count, fill = sex)) +
  geom_col() +
  coord_flip() +
  labs(x = "Cancer",
       y = "Incidences",
       title = "10 Highest Cancers in NHS Borders (2017 - 2021)",
       fill = "Sex") +
  scale_fill_manual(values = c("Male" = "#003087", "Female" = "#00A9CE" ))
```

<br>

- This plot displays the 10 most reported cancer types in NHS borders, split by gender
- It is evident that non-melanoma skin cancer is by far the most reported, with 2 other forms of skin cancers appearing in the top 10.


<br> 
<br>
<br>

To provide further insight into this I then looked into:
<br>

## **How do incidence numbers differ with age range?**

<br>

```{r}
# total rate per gender per age
hb_5yr_incidence_longer <- hb_5yr_incidence %>% 
  select(hb, cancer_site, sex, incidence_rate_age_under5:incidence_rate_age85and_over) %>% 
  filter(hb == "S08000016") %>% 
  filter(sex != "All") %>% 
    pivot_longer(cols = incidence_rate_age_under5:incidence_rate_age85and_over,names_to = "age",values_to = "rate") %>% 
  mutate(age = str_remove(string = age, pattern = "incidence_rate_age[_]*"),
         age = str_replace(string = age, pattern = "to", replacement = " - "),
         age = str_replace(string = age, pattern = "under", replacement = "<"),
         age = str_replace(string = age, pattern = "85and_over", replacement = ">85"),
        age = factor(age, levels = c("<5", "5 - 9", "10 - 14", "15 - 19", "20 - 24", "25 - 29", "30 - 34", "35 - 39", "40 - 44", "45 - 49", "50 - 54", "55 - 59", "60 - 64", "65 - 69", "70 - 74", "75 - 79", "80 - 84", ">85"))) %>% 
    group_by(age,sex) %>% 
  summarise(total_rate = sum(rate)) 

# biggest cancer per gender per age
biggest_cancer_age <- 
hb_5yr_incidence %>% 
  select(hb, cancer_site, sex, incidences_age_under5:incidences_age85and_over) %>% 
  filter(hb == "S08000016") %>% 
   filter(sex != "All") %>% 
    pivot_longer(cols = incidences_age_under5:incidences_age85and_over,names_to = "age",values_to = "rate") %>% 
  mutate(age = str_remove(string = age, pattern = "incidences_age[_]*"),
         age = str_replace(string = age, pattern = "to", replacement = " - "),
         age = str_replace(string = age, pattern = "under", replacement = "<"),
         age = str_replace(string = age, pattern = "85and_over", replacement = ">85")) %>% 
  filter(cancer_site != "All cancer types") %>% 
  group_by(age, cancer_site) %>% 
  summarise(total_incidences = sum(rate)) %>% 
 slice_max(order_by = total_incidences) %>% 
  filter(total_incidences > 1)

right_join(hb_5yr_incidence_longer, biggest_cancer_age, by ="age") %>% 
  
   mutate(age = factor(age, levels = c("<5", "5 - 9", "10 - 14", "15 - 19", "20 - 24", "25 - 29", "30 - 34", "35 - 39", "40 - 44", "45 - 49", "50 - 54", "55 - 59", "60 - 64", "65 - 69", "70 - 74", "75 - 79", "80 - 84", ">85"))) %>% 
 filter(cancer_site != "Acute lymphoblastic leukaemia") %>% 
  mutate(cancer_site = ifelse(age == "15 - 19", "Skin Cancer (Basal/Melanoma) & Thyroid", cancer_site)) %>% 
  mutate(cancer_site = ifelse(cancer_site == "Carcinoma in situ of the cervix uteri", "Cervical", cancer_site)) %>% 
  mutate(cancer_site = ifelse(cancer_site == "All brain and CNS tumours (malignant and non-malignant)", "Brain/ CNS", cancer_site)) %>% 
  mutate(cancer_site = ifelse(cancer_site == "Basal cell carcinoma of the skin", "Basal Cell Skin", cancer_site)) %>% 
  mutate(cancer_site = ifelse(cancer_site == "Bone and connective tissue", "Bone/ Tissue", cancer_site)) %>% 
  mutate(cancer_site = ifelse(cancer_site == "Colorectal cancer", "Colorectal", cancer_site)) %>% 
  mutate(cancer_site = ifelse(cancer_site == "Malignant melanoma of the skin", "M Melanoma", cancer_site)) %>% 
  mutate(cancer_site = ifelse(cancer_site == "Non-melanoma skin cancer", "Non-Melanoma Skin", cancer_site)) %>% 
  
  ggplot(aes(x = age, y = total_rate, fill = cancer_site)) + 
  geom_col() +
  coord_flip(expand = FALSE) +
  labs(y = "Rate",
       x = "Age",
  fill = "Most Reported Cancer Site in Age Group ",
  title = "Rates of Cancer per Age", 
  subtitle = "Coloured by Most Reported Cancer per Age Group")+
  

  scale_fill_manual(values = c("Non-Melanoma Skin" = "#003087", "Breast" = "#00A9CE", "Cervical" = "#009639", "Brain/ CNS" = "#ED8B00", "Leukaemias" = "#DA291C" ))
 
  

```



<br>

- This plot shows the "rate" of cancer within each age group, where each group is coloured by the most reported cancer type by that age group.
- Evidently, non-melanoma skin cancer is most prevalent in all ages above 60.
- Breast cancer is most common in patients aged between 45 to 60
- Cervical cancer is then most common in patients from 25 - 44
- In younger people, brain cancers and Leukemias are the highest reported.
_ There is a definite trend that there is a higher rate of cancer with age, with quite a large increase after the mid-60s.
- It would be useful to then split this plot by gender, as some cancers are gender specific.

<br>





<br>



------------------------------------------------------------------------

<br>

## Conclusion

<br>

- There is an increase in number of cancers report as age increases
- Skin cancers are very highly reported
- There is a low rate of incidence in groups under the age of 20
- Further insights could be gathered on how cancers differ across genders and how rates in the Borders compare to the rest of Scotland.


<br> 