---
title: "W2D1 Homework - Kirsty"
output:
  html_document:
    df_print: paged
---
--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------

# Load Libraries & Data

Libraries:

```{r}
library(tidyverse)
library(dplyr)
library(tidyr)
```

Data:

```{r}
card_base        <- read_csv("data/CardBase.csv")
customer_base    <- read_csv("data/CustomerBase.csv")
fraud_base       <- read_csv("data/FraudBase.csv")
transaction_base <- read_csv("data/TransactionBase.csv")
```

--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------

# MVP

##  Joins

### Question 1:

#### Read in all 4 credit card transaction datasets and clean column names.

```{r}
card_base        <- janitor::clean_names(card_base)
customer_base    <- janitor::clean_names(customer_base)
fraud_base       <- janitor::clean_names(fraud_base)
transaction_base <- janitor::clean_names(transaction_base)
```

### Question 2:

#### Join the data containing card details and customer details by customer id, so that all records of card details and any matching records in customer details are kept. Before you run the code, think about how many rows you expect to see after joining.

-   I would expect to see as many rows as there are in the card_base dataframe (500)

```{r}

card_customer <- card_base %>% 
  left_join(customer_base, "cust_id")

#head(card_customer)
#glimpse(card_customer)

```

### Question 3:

#### Join the data containing fraud details with transaction details so all rows of both tables are kept. What does the resulting row number tell you?

```{r}

fraud_transaction <- fraud_base %>% 
  full_join(transaction_base, "transaction_id")

#fraud_transaction

```

-   The resulting row number tells us that both tables have joined together, and any transactions that don't have a fraud flag have `NA` as their `fraud_flag` value.

### Question 4:

#### Join the data containing card details with transaction details so rows from the first which have matching ones in the second are returned, but only return rows in the first table once.

```{r}

card_transaction <- card_base %>% 
  semi_join(transaction_base, by = c("card_number" = "credit_card_id"))

#card_transaction

```
 
 -----------------------------------------------------------------------------------------------------------------------------------------------------------------------

# Tidyr
 
### Question 5

#### Read in hat_observations and separate observation into two columns, hat_colour and hat_type.

```{r}

hat_observations <- read_csv("data/hat_observations.csv")

#hat_observations

```

```{r}

hat_observations_split <- hat_observations %>% 
  separate(observation,
           into = c("hat_colour", "hat_type"),
           sep  = ",")

#hat_observations_split

```
 
### Question 6

#### Unite day, month, and year columns into a column called date using a suitable separator. Then find the date where the most berets were observed.

```{r}
hats_united <- hat_observations_split %>% 
  unite(date,
        c("day", "month", "year"),
        sep = ".")

#hats_united

hats_united %>% 
  select(date, hat_type, observation_count) %>% 
  filter(hat_type == "beret" | hat_type == "Beret") %>% 
  arrange(desc(observation_count)) %>% 
  head(1) 

```
- The most berets were observed on 18.6.2018
 
------------------------------------------------------------------------------------------------------------------------------------------------------------------------

# Extension

## Joins

### Question 1
 
#### Can you join all 4 datasets together so that you’re left with a dataset that looks like below with 109 rows and 12 columns?

- since the expected output table should only have 109 rows, the fraud dataset cannot be repeated
- `fraud_base` only shares a foreign key with `transaction_base`
- therefore to join all 4 tables, and to include all information:
  1. first fully join `customer_base` + `card_base`    <- `card_customer_full`
  2. then join `transaction_base` with this            <- `transaction_card_cust`
  3. then inner join with `fraud_base` to get the full table 
  
1.  
```{r}

card_customer_full <- card_base %>% 
  full_join(customer_base, "cust_id")

#card_customer_full

```
  

2. 
```{r}

transaction_card_cust <- card_customer_full %>% 
  full_join(transaction_base, by = c("card_number" = "credit_card_id"))

#transaction_card_cust

```


3.
```{r}

transaction_card_cust_fraud <- transaction_card_cust %>% 
  inner_join(fraud_base, by = "transaction_id")

#transaction_card_cust_fraud

```


## Tidyr

### Question 2

#### Read in exam_scores and transform it into long format with two new columns exam_question and score. Then, using separate and select, remove superfluous information from the          values in exam_question

```{r}

exam_scores <- read_csv("data/exam_scores.csv")

#exam_scores

```

```{r}
exam_scores_long <- exam_scores %>% 
  
pivot_longer(cols = starts_with("e"),
             names_to = "exam_question",
             values_to = "score")
  
#exam_scores_long  
```

```{r}

exam_scores_tidy <- exam_scores_long %>%  
  separate(exam_question,
           into = c("exam", "exam_question"),
           sep = "exam_Q") %>% 
  select("id", "exam_question", "score")

#exam_scores_tidy

```


 
 