---
title: "W6D2_homework_kirsty"
output: html_notebook
---

# MVP

## Q1: Load the data, clean_names() and explore it.

```{r}
# library(tidyverse)
# library(janitor)
```

```{r}
mobiles <- read_csv("data/20190928-items.csv")

# mobiles

mobiles <- clean_names(mobiles)
```





## Q2: Find the brand with the highest numbers of phones in the dataset.
```{r}
mobiles %>% 
  group_by(brand) %>% 
  count() %>% 
  arrange(desc(n)) %>% 
  head(2) # ensure that the top two counts aren't the same
```
- The brand with the highest number of phones is Samsung. 





## Q3: For your top brand, plot the distribution of phone ratings as a probability density, overlaying a fitted normal distribution. Do you think the normal distribution provides a good model of these ratings?

```{r}
# Create table detailing total ratings, mean, and SD
samsung_stats <- mobiles %>% 
  filter(brand == "Samsung") %>% 
  summarise(num = n(),
            mean = mean(rating),
            sd = sd(rating))
```


```{r}
mobiles %>% 
  filter(brand == "Samsung") %>% 
  ggplot(aes(x = rating)) +
  geom_histogram(aes(y = after_stat(density)), colour = "white") +
  stat_function(
    fun = dnorm,
    args = list(
      mean = samsung_stats$mean,
      sd = samsung_stats$sd
    ),
    col = "red"
  )

```

The normal distribution is not a good model for these ratings as the histogram is not unimodal and has various peaks throughout, including peaks at both sides of the graph.



## Q4: Use these values to calculate the proportions of ratings within one-, two- and three standard deviations of the mean. Compare these proportions with those you expect for a normal distribution.

within 1 SD = rating is between (mean - 1SD) and (mean + 1SD), 
i.e. more than (mean - SD) and less than (mean + SD)
= >(3.5733 - 0.6847859) & <(3.5733 + 0.6847859)
= between 2.888514 - 4.258086

                                                                    
```{r}
mobiles %>% 
  filter(brand == "Samsung") %>% 
  filter(rating >= samsung_stats$mean - samsung_stats$sd) %>% 
  filter(rating <= samsung_stats$mean + samsung_stats$sd) 
```
315/397 = 79.34% of ratings are within 1 SD of the mean

2SD: 

```{r}
mobiles %>% 
  filter(brand == "Samsung") %>% 
  filter(rating >= samsung_stats$mean - (2*(samsung_stats$sd))) %>% 
  filter(rating <= samsung_stats$mean + (2*(samsung_stats$sd)))
```
365/397 = 91.94% of ratings are within 2 SD of the mean

3SD:

```{r}
mobiles %>% 
  filter(brand == "Samsung") %>% 
  filter(rating >= samsung_stats$mean - (3*(samsung_stats$sd))) %>% 
  filter(rating <= samsung_stats$mean + (3*(samsung_stats$sd)))
```
389/397 = 97.98% of ratings are within 3 SD of the mean

Could also do this using the `scale()` function to create Z scores

```{r}
# create table with z scores:

mobiles_scaled <- mobiles %>% 
  filter(brand == "Samsung") %>% 
  mutate(z_rating = scale(rating), .after = rating)
```

```{r}
# filter where Z is between +/- 1, so therefore +/- 1 SD away from the mean
mobiles_scaled %>% 
  filter(between(z_rating, left = -1, right = 1))
```
315/397 = 79.34% of ratings are within 1 SD of the mean

```{r}
mobiles_scaled %>% 
  filter(between(z_rating, left = -2, right = 2))
```
365/397 = 91.94% of ratings are within 2 SD of the mean

```{r}
mobiles_scaled %>% 
  filter(between(z_rating, left = -3, right = 3))
```

389/397 = 97.98% of ratings are within 3 SD of the mean

For a normal distribution, these proportions should be: 
68%, 95%, & 99.7%. As the proportions of this distribution are different, it is therefore not a normal distribution and is likely to be skewed. 
