---
title: "W6D3_homework_kirsty"
output: html_notebook
editor_options: 
  markdown: 
    wrap: 72
---

```{r}
library(tidyverse)
library(janitor)
library(infer)
```

# MVP

## Q1: Load the data, clean_names(), and familiarise yourself with it.

```{r}
ames <- read_csv("data/ames.csv") %>%  clean_names()

ames
```

## Q2: Investigate the distribution of lot_area. Is the distribution roughly normal? If not, what problems do you find?

```{r}
ames %>% 
  ggplot(aes(x = lot_area)) +
  geom_histogram()
```
The shape of the plot is unimodal which suggests it is fairly normal however it is skewed to the right because of the right hand tail and outlier.


## Q3: Compute and visualise a bootstrap sampling distribution for the mean(lot_area) of the sold houses.

```{r}
# Compute using `infer`, with 5000 replicates

ames_sample <- ames %>% 
  specify(response = lot_area) %>% 
  generate(reps = 5000, type = "bootstrap") %>% 
  calculate(stat = "mean")

```

```{r}
# visualise using `infer`

ames_sample %>% 
  visualise()

```

## Q4: Use your bootstrap distribution to calculate a 95% CI for mean(lot_area), and visualise it on the distribution

```{r}
# Calculate 95% CI

ames_ci_95 <- ames_sample %>% 
  get_confidence_interval(level = 0.95, type = "percentile")

```

```{r}
# add CI to visualisation

ames_sample %>% 
  visualise() +
  shade_confidence_interval(endpoints = ames_ci_95)

```

## Q5: You would like to know the mean(lot_area) of the sold houses with higher confidence. Calculate the 99%

CI for this variable (you can re-use your bootstrap distribution from
above). Is it narrower or broader than the 95% CI? Does that make sense?

```{r}
# Calculate 99% CI

ames_ci_99 <- ames_sample %>% 
  get_confidence_interval(level = 0.99, type = "percentile")

ames_ci_99

# for reference the 95% CI:

ames_ci_95 

```

The 99% CI is broader than the 95%, this makes sense as for the 99% CI
you are saying that 99% of samples are within this CI, so the limits of
this CI must be broader to "catch" more samples.

## Q6: Calculate the point estimate of the mean(lot_area)

```{r}
ames_sample %>% 
  summarise(point_est = mean(stat))
```

# Extension

Calculate a point estimate and 95% CI for the proportion of houses in
the data built before 1920. Does the number of reps you use matter?
[Investigate reps from 200 up to 50000, memory of your laptop
permitting].

For 200 reps:

```{r}
# create a bootstrap sampling distribution

b_1920_200_sample <- ames %>% 
  # since "success" needs a string, add a column that says "yes" if built <1920
  mutate(before_1920 = ifelse(year_built <1920, "yes", "no"), 
         .after = year_built) %>% 
   # now look for "yes" responses in this new column
  specify(response = before_1920, success = "yes") %>%
  generate(reps = 200, type = "bootstrap") %>% 
  calculate(stat = "prop")

```

```{r}
# now get the 95% CI
b1920_200_ci_95 <- b_1920_200_sample %>% 
  get_confidence_interval(level = 0.95, type = "percentile")
```

```{r}
# now get the point estimate
b_1920_200_sample %>% 
  summarise(point_est_200 = mean(stat))
```

```{r}
b_1920_200_sample %>% 
  visualise() +
  shade_confidence_interval(endpoints = b1920_200_ci_95)
```

For 1000 reps:

```{r}

b_1920_1k_sample <- ames %>% 
  mutate(before_1920 = ifelse(year_built <1920, "yes", "no"), .after = year_built) %>% 
  specify(response = before_1920, success = "yes") %>% 
  generate(reps = 1000, type = "bootstrap") %>% 
  calculate(stat = "prop")

b_1920_1k_ci_95 <- b_1920_1k_sample %>% 
  get_confidence_interval(level = 0.95, type = "percentile")
b_1920_1k_ci_95



point_est_1k <- b_1920_1k_sample %>% 
  summarise(point_est_1k = mean(stat))
point_est_1k

```

For 10K reps:

```{r}
b_1920_10k_sample <- ames %>%
  mutate(before_1920 = ifelse(year_built <1920, "yes", "no"), .after = year_built) %>%
  specify(response = before_1920, success = "yes") %>%
  generate(reps = 10000, type = "bootstrap") %>%
  calculate(stat = "prop")

b_1920_10k_ci_95 <- b_1920_10k_sample %>% 
  get_confidence_interval(level = 0.95, type = "percentile")
b_1920_10k_ci_95



point_est_10k <- b_1920_10k_sample %>% 
  summarise(point_est_10k = mean(stat))
point_est_10k

```

For 50K reps:

```{r}
 b_1920_50k_sample <- ames %>%
  mutate(before_1920 = ifelse(year_built <1920, "yes", "no"), .after = year_built) %>%
  specify(response = before_1920, success = "yes") %>%
  generate(reps = 50000, type = "bootstrap") %>%
  calculate(stat = "prop")

b_1920_50k_ci_95 <- b_1920_50k_sample %>% 
  get_confidence_interval(level = 0.95, type = "percentile")
b_1920_50k_ci_95



point_est_50k <- b_1920_50k_sample %>% 
  summarise(point_est_50k = mean(stat))
point_est_50k
```

| Number of Reps | Point Estimate |  Lower CI  |  Upper CI   |
|:--------------:|:--------------:|:----------:|:-----------:|
|      200       |   0.06014164   | 0.0518686  | 0.06794369  |
|       1K       |   0.05996758   | 0.05187713 | 0.06825939  |
|      10K       |   0.06010512   | 0.05153584 | 0.06894198  |
|      50K       |   0.06004466   | 0.05153584 | 0.06860068  |

As rep number increases the CI should get wider and point estimate should get more accurate.