---
title: "W5D1_homework_kirsty"
output: html_notebook
---

# Load Libraries & Data

```{r}
library(tidyverse)
library(CodeClanData)
```

```{r}
revenue_breakdown <- qb_revenue_breakdown
monthly_sales     <- qb_monthly_sales
competitors       <- qb_competitors
device_data       <- qb_device_data
```

------------------------------------------------------------------------

# MVP

## Part 1

#### Q1:

#### Take the data in the dataset qb_revenue_breakdown and make a stacked bar chart showing the sources of revenue across the two years in the dataset.

A:

```{r}
revenue_breakdown %>% 
  ggplot() +
  geom_bar(aes(x = Year, 
               y = Revenue, 
               fill = Product), 
           stat = "identity")
```

#### Q2:

#### Make a line chart showing monthly sales of the "flagship product" - the 'KwikBit Unit 1' - over the last year. This data is inside qb_monthly_sales, but you will need to use filter() or subset() (base R).

```{r}
monthly_sales %>% 
  filter(Cashflow == "Kwikbit Unit 1 Sales") %>% # filter for Kwikbit only
  ggplot() +
  geom_line(aes(x = Date, y = Amount))
```

#### Q3:

#### Make a line chart showing monthly revenue and costs over the last year. This data is also in qb_monthly_sales.

```{r}
monthly_sales %>% 
  filter(Cashflow != "Kwikbit Unit 1 Sales") %>% # filter out Kwikbit
  ggplot() +
  geom_line(aes(x = Date,
                y = Amount, 
                group = Cashflow))
```

#### Q4:

#### Show annual sales of personal fitness trackers over the last 5 years broken down by company as a ribbon plot (use geom_area). This data is in qb_competitors.

```{r}
competitors %>% 
  ggplot() +
  geom_area(aes(x = Year, 
                y = Revenue, 
                fill = Company))
```

#### Q5:

#### Now show the sales from the four competitors as a line graph. Include an extra layer that shows the data points used to make the lines.

```{r}
competitors %>% 
  ggplot() +
  geom_line(aes(x = Year, 
                y = Revenue, 
                group = Company, 
                colour = Company)) +
  geom_point(aes(x = Year, # new layer
                 y = Revenue, 
                 colour = Company)) 

```

#### Q6:

#### Now the company wants to compare the number of steps that their device counts vs. their competitors. Make a line graph of the number of steps throughout time, and use faceting to compare between companies and people. The data for this is in qb_device_data.

```{r}
device_data %>% 
  ggplot() +
  geom_line(aes(x = hours, 
                y = counts, 
                group = id, 
                colour = device), 
            stat = "identity") +
  facet_grid(device ~ id)


```

## Part 2:

I will copy the same base code used in Part 1

#### Q1: Sources of Revenue against Year

```{r}
col_scheme <- c("#E89FE9", "#50434F", "#B6A7B5", "#F9A472", "#BD7040")

revenue_breakdown %>% 
  ggplot() +
  geom_bar(aes(x = Year, 
               y = Revenue, 
               fill = Product), 
           stat = "identity") +
  ggtitle("Revenue per Year by Product") +
  theme_light(base_size = 12) +
  scale_fill_manual(values = col_scheme) + # use given scheme colours to fill
  theme(axis.title = element_text(size = 10)) +
  theme(title = element_text(size = 14))
```

#### Q2: Monthly Sales of KwikBit

```{r}

monthly_sales %>% 
  filter(Cashflow == "Kwikbit Unit 1 Sales") %>% 
  ggplot() +
  geom_line(aes(x = Date, 
                y = Amount), 
            colour = "#E89FE9") +
  xlab("\nMonth") +
  ylab("Revenue\n") +
  scale_x_date(breaks = "1 month", date_labels = "%b %y") +
  ggtitle("Monthly Sales of Kwikbit") +
  theme_light(base_size = 12) +
  theme(axis.title = element_text(size = 10)) +
  theme(title = element_text(size = 14))
```

#### Q3: Monthly Revenue vs Costs

```{r}
monthly_sales %>% 
  filter(Cashflow != "Kwikbit Unit 1 Sales") %>% 
  ggplot() +
  geom_line(aes(x = Date, 
                y = Amount, 
                group = Cashflow, 
                colour = Cashflow), 
            position = "dodge") +
  scale_colour_manual(values = c("#BD7040","#50434F")) +
  xlab("\nMonth") +
  ylab("Amount\n") +
  scale_x_date(breaks = "1 month", # simplify date to month/year
               date_labels = "%b %y") +
  scale_y_continuous(
    breaks = c(1500, 2000, 2500, 3000), # scale Y axis
    limits = c(1500, 3000)) +
  ggtitle("Monthly Revenue vs Costs") +
  theme_light(base_size = 12) +
  theme(axis.title = element_text(size = 10)) +
  theme(title = element_text(size = 14)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5))



```

#### Q4 : Annual Sales by Company

```{r}
competitors %>% 
  mutate(sales_millions = Revenue / 1000000) %>% # scale Y axis
  ggplot() +
  geom_area(aes(x = Year, 
                y = sales_millions, 
                fill = Company)) +
  scale_fill_manual(values = col_scheme) +
  xlab("\nYear") +
  ylab("Revenue (millions)\n") +
  scale_y_continuous(labels = scales::comma, 
                     limits = c(0,7)) +
  ggtitle("Annual Sales by Company") +
  theme_light(base_size = 12) +
  theme(axis.title = element_text(size = 10)) +
  theme(title = element_text(size = 14)) 
 
```

#### Q5 : Annual Sales by Company

```{r}
competitors %>% 
  ggplot() +
  geom_line(aes(x = Year, 
                y = Revenue, 
                group = Company, 
                colour = Company)) +
  geom_point(aes(x = Year, 
                 y = Revenue, 
                 colour = Company)) +
  scale_colour_manual(values = col_scheme) +
  xlab("\nYear") +
  ylab("Revenue\n") +
  scale_y_continuous(labels = scales::comma) +
  ggtitle("Annual Sales by Company") +
  theme_light(base_size = 12) +
  theme(axis.title = element_text(size = 10)) +
  theme(title = element_text(size = 14)) 
```

#### Q6: Step Count per Competitor and Person

```{r}
device_data %>% 
  mutate(count_thousands = counts / 1000) %>% # scale Y axis
  ggplot() +
  geom_line(aes(x = hours, 
                y = count_thousands, 
                group = id, 
                colour = device), 
            stat = "identity") +
  scale_colour_manual(values = col_scheme) +
  facet_grid(device ~ id ) +
  xlab("\nHour") +
  ylab("Steps (thousands)\n") +
  ggtitle("Step Count by Hour per Person and Competitor") +
  labs(colour = "Device") +
  theme_light(base_size = 12) +
  theme(axis.title = element_text(size = 10)) +
  theme(title = element_text(size = 14)) 
  
```
