---
title: "Decision Trees Homework"
output:
  html_document:
    df_print: paged
---

**Load Libraries**

<br>

```{r message=FALSE, warning=FALSE}
library(rpart)
library(rpart.plot)
library(tidyverse)
library(GGally)
library(janitor)
library(modelr)
library(yardstick)

titanic_set <- read_csv('data/titanic_decision_tree_data.csv')

shuffle_index <- sample(1:nrow(titanic_set))

# shuffle the data so class order isn't in order - need this for training/testing split later on 
titanic_set <- titanic_set[shuffle_index, ]
```

<br>

# MVP

<br>

## 1. Clean Data

```{r}
(titanic_set_clean <- titanic_set %>%
  drop_na(survived) %>%  # drops 418 rows
  mutate(sex = as.factor(sex),
         survived = factor(survived, levels = c(0,1), labels = c("Died", "Survived")),
         pclass = as.factor(pclass),
         embarked = as.factor(embarked)) %>% 
  mutate(age_status = if_else(age <= 16, "child", "adult")) %>%
    select(-...1, -passenger_id, -name, -ticket, -fare, -cabin, -age) %>% 
   na.omit() # drops 179 rows
  
)
```

<br>

## 2. General plots

```{r message=FALSE, warning=FALSE}
titanic_set_clean %>% 
ggpairs(progress = FALSE)
```

From the GGpairs plot, I would possibly expect sex, class, embarkment port, siblings/spouses count, and age to have an impact.

<br>

```{r}
titanic_set_clean %>% 
  group_by(survived) %>% 
  count() %>% 
  mutate(percentage = round(((n/712)*100),digits = 0))

```

60% of the passengers in the dataset survived, whereas 40% died.

<br>

```{r}
# check sex ratio

titanic_set_clean %>% 
  group_by(sex) %>% 
  count() %>% 
  mutate(percentage = round(((n/712)*100),digits = 0)) # 712 passengers in cleaned data set
```
36% of the passengers in the dataset were female, 64% were male

<br>

```{r}
# check classes and their counts

titanic_set_clean %>% 
  group_by(pclass) %>% 
  count() %>% 
  mutate(percentage = round(((n/712)*100),digits = 0))
```
26% of passengers were in 1st class, 24% in 2nd, and 50% in third

<br>

```{r}
titanic_set_clean %>% 
  group_by(embarked) %>% 
  count() %>% 
  mutate(percentage = round(((n/712)*100),digits = 0))
```
78% of the passengers embarked at Southampton, 18% at Cherbourg, and 4% at Queenstown

<br>

```{r}
titanic_set_clean %>% 
  group_by(sib_sp) %>% 
  count() %>% 
  mutate(percentage = round(((n/712)*100),digits = 0))
```
66% of passengers had no spouses or siblings on board, whereas 34% had one or more.

<br>

```{r}
titanic_set_clean %>% 
  group_by(age_status) %>% 
  count() %>% 
  mutate(percentage = round(((n/712)*100),digits = 0))
```
86% of passengers were adults whereas 14% were children.

<br>
<br>

## 3. Build Model

```{r}
#80% of data will be used in training, 20% in test

# get total row count
n_data <- nrow(titanic_set_clean)

# create test sample index
test_index <- sample(1:n_data, size = n_data*0.2)

# create test set
titanic_test <- slice(titanic_set_clean, test_index)

# create training set
titanic_train <- slice(titanic_set_clean, -test_index)

```


```{r}
# check proportions of survived/died are similar across test and train sets

titanic_test %>% 
  tabyl(survived)

titanic_train %>% 
  tabyl(survived)
```
- The proportions look similar in both sets

<br>

## 4. Build tree model 

```{r}
titanic_fit <- rpart(
  formula = survived ~ .,
  data = titanic_train,
  method = 'class' # as this is used for categorical variables
)

rpart.plot(titanic_fit,
           yesno = 2,
           fallen.leaves = TRUE,
           faclen = 6,
           digits = 4)
```

<br>

## 5. Analyse Tree Plot

- The model has picked sex, class and number of siblings/spouses on board as the most informative predictors.
- The root node tells us that when taking all data into account the most likely result is "died", as there is a low (41%) probability of surviving
- It also tells us that 100% of the data passed through this node (as expected with the root node)
- The root node then leads to the first classification category related to gender: 
    - If male, then there is a 20% chance of survival 
    - You might expect this as women and children were prioritised on the life boats
    - 64% of the data passed through this node - as there was only one category leading to this node then 64% of the passengers in the data must have been male
    - If female, then we can see at this level the chance of surviving is 76%, leading to an overall likelihood of survival. This node tells us that 36% of the passengers were        female
- This `female` branch is then further classified into the class of the passenger:
  - If the passengers class was 3, then at this level the passenger is not likely to have survived, as they had a probability of 48% survival. 14% of the passengers were female     and of class 3
- This `female, class 3` branch is further split into the number of a passenger's relatives on board:
  - If the passenger was female and of class 3 **and** had other family (siblings or spouses) then they had a 35% chance of survival so most likely died
      - this may be from passengers wanting to be with their relatives and lowering their chance of survival
  - 6% of passengers were in this category
  - If the passenger did not have other family (7.54% of female class 3 passengers) then it is 60% likely that they survived
- Returning to the females that were *not* class 3, it is highly likely that they survived (94%). 22% of the data got to this stage

<br>

## 6. Test and add predictions

```{r}
titanic_test_predictions <- titanic_test %>% 
  add_predictions(titanic_fit, type = "class")
```

```{r}
titanic_test_predictions %>% 
  select(sex, pclass, sib_sp, survived, pred)
```

From a quick glance at the data, the predictions look to be quite accurate!
Create a confusion matrix:


```{r}
(conf_mat <- titanic_test_predictions %>% 
  conf_mat(truth = survived,
           estimate = pred)
)
```
The column on the left is the predicted values, and the actual values are on the top row
Therefore 76 passengers were correctly predicted to die, and 36 were correctly predicted to survive.
This is a lot higher than the 10 passengers that were predicted to survive but actually died, and the 20 passengers that were predicted to die but actually survived.
This can also be represented using the `accuracy` function:

```{r}
(accuracy <- titanic_test_predictions %>% 
   accuracy(truth = survived,
            estimate = pred))
```
This tells us that there is a 79% chance of predicting if a passenger survived using this model.
