---
title: "SLR_homework"
output:
  html_document:
    df_print: paged
---


```{r}
library(tidyverse)
library(broom)
library(ggfortify)
```


# MVP

## Load Data

```{r}
project <- read_csv("data/project_management.csv")

project
```

<br>

## Plot Data 

```{r}
project %>% 
  ggplot(aes(x = estimated_length, y = actual_length)) +
  geom_point()
```
<br>

## Correlation Coefficient
```{r}
project %>% 
  summarise(correlation = cor(x = estimated_length, y =actual_length))
```
As the correlation coefficient is >0.5, this is therefore a strong positive correlation.

<br>

## Simple Linear Regression 

```{r}
(project_model <- lm(
  formula = actual_length ~ estimated_length,
  data = project
))
```

<br>

## Regression Coefficient and r squared


A 1 unit increase in the explanatory variable value (estimated_length) changes the outcome variable value (actual_length) by an amount equal to b1 (the slope).

In this case a 1 day increase in actual length changes the predicted length by 1.223 days

```{r}
glance(project_model) %>%  janitor::clean_names()
```
As this is a linear regression on a single variable, then $r^2$ is the square of the correlation coefficient (r).  


$r^2$ = 0.8046597	* 0.8046597	= 0.6474772

Since $r^2$ is the fraction of the variation size in the outcome variable that can be explained by the amount of variation in the explanatory variable, in this example 64.75% of the variation in estimated number of days can be explained by the variation in the actual number of days.
The closer to 1 the $r^2$ value is, the better the lines fit to the model is, an r squared value of 1 indicates that all of the variation in the outcome variable is explained by the variation in the explanatory variable.

<br>

## Reviewing Statistical Difference

Check diagnostic plots to gauge P value's reliability:

```{r}
autoplot(project_model)
```

<br>

Residuals vs Fitted:
- There does seem to be a bit of a pattern with the points, they don't appear to be randomly scattered
- Therefore would question if the residuals are independent of each other
- However, the smoothed line does stay relatively close to 0 but drops off at the end, because of this I would say that the residuals are independent

<br>

Normal Q-Q:
- The points all lie relatively close to the line, with a very small amount of variation in the middle
- Therefore would say that the residuals are normally distributed

<br>

Scale Location:
- No evidence of heteroscedasticity suggests that the variance is constant

Given that these plots mostly hold up against their requirements, I would say that the p_value is reliable.


```{r}
tidy(project_model) %>% janitor::clean_names()
```
The p value of the regression coefficient (estimated_length) is 1.911284e-12, this is less than a typical significance level of 0.05 or 0.01, so therefore the evidence suggests that the relationships are statistically different.




