---
title: "W6D1_homework_kirsty"
output: html_notebook
---

```{r}
library(tidyverse)
library(gtools)
library(janitor)
source("data/prob.R")
```

# MVP

## 1.1 The Titanic

### Q1: Calculate the conditional probabilities prob(survived∣crew) and prob(survived∣passenger)

Have a look at the distinct values of the class variable, this will tell you whether each individual was a passenger (1st, 2nd or 3rd class) or a member of the crew (treat restaurant staff as crew)


#### prob(survived∣crew)
```{r}
titanic <- read_csv("data/titanic.csv")

titanic %>% 
  distinct(class)
```


```{r}

titanic <- titanic %>% 
  mutate(crew_passenger = ifelse(str_detect(class, "[0-9]+"), "passenger", "crew"))
  
```

Using janitor::tabyl adorned with percentages
```{r}
titanic_table <- titanic %>% 
  tabyl(crew_passenger, survived) %>% 
  adorn_totals(where = c("row", "col"))

titanic_table
```

```{r}
titanic_table %>% 
  adorn_percentages(denominator = "all")

```

probability of being crew = 0.4033
probability of surviving = 0.3222
probability of being crew and surviving 0.0956

prob(survived|crew)=prob(survived & crew) / prob(crew)
= 0.0956 / 0.4033 = 0.237


##### Using dplyr
```{r}
titanic %>% 
  filter(crew_passenger == "crew") %>% 
  summarise(probability = sum(survived == "yes") / n())
```
####  prob(survived∣passenger)

probability of being passenger = 0.5967
probability of being passenger and surviving 0.2265
prob(survived|passenger)=prob(survived & passenger) / prob(passenger)
= 0.2266 / 0.5967 =  0.3797

##### Using dplyr

```{r}
titanic %>% 
  filter(crew_passenger == "passenger") %>% 
  summarise(probability = sum(survived == "yes") / n())
```

## 1.2: Probability Consolidation Questions

### Q1: If I roll a six-sided dice once, what is the probability of getting a number less than 5?

p(event) = n(outcomes in event) / n(total outcomes in sample space)


outcomes in event of roll <5: {1, 2, 3, 4} = 4
total outcomes of rolling a six sided dice once: {1, 2, 3, 4, 5, 6} = 6

p(<5) = 4/6 = 0.667

```{r}
4/6
```





### Q2: How many outcomes are there in the sample space of an experiment in which we toss three dice?


There are 6 outcomes in each die, 3 times = 6^3 = 216


```{r}
# n = how many options there are (6 in a die roll)
# r = how many things are being chosen (3 rolls)
permutations(n =6,
             r = 3,
             repeats.allowed = TRUE) %>% # assuming that "1,2,2" is counted as different combination from "2,2,1"
  as_tibble()
```



### Q3: In an experiment where we roll four dice:
How many outcomes are in the event where the total of the four dice equals 5, 6 or 7?
what is the probability of this event?
what assumptions do we apply in answering part ii?

```{r}
# create total column 
four_dice <- permutations(n = 6,
                          r = 4,
                          repeats.allowed = TRUE) %>% 
             as.tibble() %>% 
             mutate(total = rowSums(.))

four_dice
```
There are 1296 total possible outcomes.

```{r} 
# filter where total is 5, 6, or 7
four_dice %>% 
  filter(total == (5:7)) 
```

i .There are 9 possible combinations where the total is between 5 and 7.

ii. The probability of this event is:
  p(event) = n(outcomes in event) / n(total outcomes in sample space)

  =

  p(roll 5:7) = 9 / 1296 = 0.006944444

iii. I assume that repeats are allowed, e.g. 1,2,1,3 is treated as a different combination from 2, 1, 1, 3 etc.




### Q4: Five chairs are placed in a row in a room, along with five people of differing height. The people sit down randomly on the chairs. What is the probability that they sit in order of decreasing height, left to right?
```{r}
permutations(n = 5, # 5 chairs
             r = 5, # 5 people
             repeats.allowed = FALSE) %>% # they can't sit in >1 chair
  as.tibble()
```

There are 120 possible combinations of how they could sit.
They could only sit in height order in 1 combination. 
Therefore p(event) = n(outcomes in event) / n(total outcomes in sample space)
p(height order) = 1/ 120 0.008333333

### Q5: I roll six dice - 
i. what is the probability I see the numbers 6, 5, 4, 3, 2 and 1 in order on the dice, left to right?
ii. what is the probability I see one each of the numbers 6, 5, 4, 3, 2 and 1 on the dice regardless of order?

i. 
```{r}
permutations(n = 6,
             r = 6,
             repeats.allowed = TRUE) %>%  # the numbers rolled are independent of each other
  as.tibble()
```

There are 46656 possible combinations, of this there is only one opportunity for the dice to land in order
Therefore p(event) = n(outcomes in event) / n(total outcomes in sample space)
p(ordered roll) = 1 / 46656 = 2.143347e-05

ii. 
```{r}
six_die <-
permutations(n = 6,
             r = 6,
             repeats.allowed = TRUE) %>%  # the numbers rolled are independent of each other
  as.tibble() 
```


```{r}
six_die %>% 
  filter(isin(., c(6,5,4,3,2,1), ordered = FALSE)) %>% 
  as.tibble()
```
ii. There are 720 possible combinations of 1:6
p(event) = n(outcomes in event) / n(total outcomes in sample space)
p(1:6) = 720 / 46656 = 0.0154321


### Q6: If I draw a single card from a 52-card deck, and it is black, what is the probability that its suit is spades?



```{r}
cards(jokers = FALSE) %>% 
  mutate(colour = if_else(suit == "Club"|suit == "Spade", "black", "red")) %>% 
  filter(colour == "black") %>% 
  summarise(probability = sum(suit == "Spade")/n())
```



