---
title: "Homework_Weekend1_Kirsty"
output:
  html_document:
    df_print: paged
editor_options: 
  markdown: 
    wrap: 72
---

------------------------------------------------------------------------

## Load Packages & Data

```{r}
library(tidyverse)
library(dplyr)
library(stringr)
library(knitr)
```

```{r}
books <- read_csv("data/books.csv")
```

------------------------------------------------------------------------

## Exploring the Data

```{r}
glimpse(books)
```

-   There are 11123 rows and 13 columns
-   Variables are either of the type double or character

The variable names are:

```{r}

names(books)

```

------------------------------------------------------------------------

## Cleaning the Data

-   There are some columns that don't follow the snake_case style
    -   Fix this using `janitor`, and save data under a new name

```{r}

books_cleaned <- janitor::clean_names(books)

#glimpse(books)

```

-   Check for `NA` values:

```{r}

books_cleaned %>% 
  summarise(across(rowid:publisher, ~sum(is.na(.x))))

```

There are no `NA` values!

-   Noted issues:
    -   It appears that there are some publishers and some titles that
        are likely to be the same but have been entered wrongly.

------------------------------------------------------------------------

## Analysing the Data

### Publisher:

#### What publisher has the most entries in the table?

```{r}

most_titled_pub <- books_cleaned %>% 
  group_by(publisher) %>% 
  summarise(title_count = (n_distinct(title))) %>% 
  arrange(desc(title_count)) %>% 
  head(1) 

most_titled_pub

```

-   The publisher with the most titles in this data set is called
    "Vintage".

#### What publisher has the best overall rating?

```{r}

avg_overall_rating <- books_cleaned %>% 
  group_by(publisher) %>% 
  mutate(avg_overall_rating = (mean(average_rating))) %>% 
  arrange(desc(avg_overall_rating)) %>% 
  ungroup() %>% 
  select(publisher, avg_overall_rating)


avg_overall_rating

```

-   There are 12 publishers all with an average overall rating of 5.
    However, all of these publishers all only have 1 book that is rated
    so therefore the results could be scewed.

#### Are there any publishers that have published in \>1 language?

```{r}
books_cleaned %>% 
  group_by(publisher) %>% 
  summarise(language_count = n_distinct(language_code)) %>% 
  arrange(desc(language_count)) %>% 
  filter(language_count >1)
```

-   There are 507 publishers that have published in \>1 language code,
    "Oxford University Press USA" have published in the most languages,
    which I would assume to be English, US English, British English,
    Spanish, Italian and greek:

```{r}
books_cleaned %>% 
  select(publisher, language_code) %>% 
  filter(publisher == "Oxford University Press  USA") %>% 
  distinct(language_code, publisher) #to remove duplicate entries
```

### Authors:

-   There are a lot of authors that are entered with their illustrator/
    co-author, to look at lead authors only:

```{r}
books_cleaned <- books_cleaned %>% 
  mutate(lead_author = gsub("/..*", "", authors), 
         .after = authors) %>% 
  arrange(desc(authors))

books_cleaned
```

#### What author has the highest overall rating?

```{r}
max_rated_author <- books_cleaned %>% 
  group_by(lead_author) %>% 
  mutate(avg_rating = (mean(average_rating))) %>% 
  arrange(desc(avg_rating)) %>% 
  select(lead_author, average_rating)

max_rated_author
```

-   There are 17 authors with an average rating of 5.00.

#### What author has the most titles listed?

```{r}
most_titled_author <- books_cleaned %>% 
  group_by(lead_author) %>% 
  summarise(title_count_author = (n_distinct(title))) %>% 
  arrange(desc(title_count_author)) %>% 
 head(1) 

most_titled_author
```

-   The author with the most titles in this data is Stephen King, who
    has 63 books listed in the table:

```{r}
books_cleaned %>% 
  select(title, lead_author) %>% 
  filter(lead_author == "Stephen King") %>% 
  distinct(title, lead_author) #to remove duplicate entries
```

#### Are there any authors that have published in \>1 language?

```{r}
books_cleaned %>% 
  group_by(lead_author) %>% 
  summarise(language_count = n_distinct(language_code)) %>% 
  arrange(desc(language_count)) %>% 
  filter(language_count >1)
```

-   There are 721 authors that have published in \>1 language, with J.K
    Rowling being the most at 7 languages. I assume these languages are
    Turkish, German, Latvian, English, Chinese, Spanish and Gaelic(?)

```{r}
books_cleaned %>% 
  select(lead_author, language_code) %>% 
  filter(lead_author == "J.K. Rowling") %>% 
  distinct(language_code, lead_author) #to remove duplicate entries
```

#### Are there any authors that have \>1 publisher?

```{r}
author_max_publishers <- books_cleaned %>% 
  group_by(lead_author) %>% 
  summarise(publisher_count = n_distinct(publisher)) %>% 
  arrange(desc(publisher_count)) %>% 
  head(1) %>%  
  pull(lead_author)

author_max_publishers
```

-   Stephen King has the most publishers (46).

## Annual Data:

-   NB: The data is entered in the format DD/MM/YY. As the date range
    for this data set is long, I will group the data for each year using
    the `str_sub()` function:

```{r}
books_years <- books_cleaned %>% 
  mutate(publication_year = str_sub(publication_date, -4),
         .after = publication_date) %>% 
  arrange(publication_year) 

books_years
```

#### What was the best book per year?

```{r}
books_years %>% 
  mutate(publication_year = as.integer(publication_year)) %>% 
  select("publication_year", "title", "average_rating") %>% 
  arrange(publication_year, desc(average_rating)) %>% 
  group_by(publication_year) %>% 
  slice_max(average_rating, n=1)  
```

-   The table above shows the best book per year based on average
    review. For any years where there are multiple books of the same
    highest rating, they are all included.

#### How many books were rated in each year?

```{r}
books_years %>% 
  group_by(publication_year) %>% 
  summarise(books_per_year = n_distinct(title)) %>% 
  arrange(publication_year) #%>% 
  #arrange(desc(books_per_year)) #to check year with most ratings
```

-   The table above shows the number of books per year.
-   The year with the highest number of books was 2006 with 1674 books
    reviewed.

------------------------------------------------------------------------

## An attempt to solve the duplicate publisher issue!

-   It appears that there are some publishers that have the same title
    both with slightly different names.
-   According to the data card, the ISBN13 number is a unique
    identifying code for each book. Each part of the code has a
    different meaning:

```{r, out.width = "75%"}
url_ISBN13 <- "https://blog-cdn.reedsy.com/directories/admin/attachments/large_How-to-read-an-ISBN-2d4aca.jpg"
include_graphics(url_ISBN13)
```

-   Since the 5-7th characters denote the publisher, we could use these
    to identify the publisher for each book. This would negate the issue
    of having duplicated publisher entries with slightly different
    names.

-   First, confirm that all ISBN-13 codes are actually 13 characters
    long, so therefore we can assume the 5-7th characters represent the
    publisher:

```{r}
books_cleaned %>% 
  
  mutate(isbn13_count = nchar(isbn13)) %>% 
  filter(isbn13_count != 13)
```

-   Therefore all ISBN13 codes are 13 characters long, so assume they
    follow the described format.

-   Next, split ISBN13 codes up, to get 5-7th character, and save this
    as a new variable called `isbn13_publisher`

```{r}
books_isbn_pub <- books_cleaned %>% 
  mutate(isbn13_publisher = substr(isbn13, 5, 7),
         .after = publisher) %>% 
  arrange(publisher)
books_isbn_pub
```

-   Now check how many publishers each isbn13 publisher code has,
    hopefully only 1!

```{r}
books_isbn13_pub_grouped <- books_isbn_pub %>% 
  group_by(publisher) %>% 
  summarise(count_isbn_pub = n_distinct(isbn13_publisher)) %>% 
  arrange(desc(count_isbn_pub)) %>% 
  filter(count_isbn_pub  >1)

books_isbn13_pub_grouped

```

-   There are 444 publishers that have more than one identifier in the
    ISBN13 codes, so unfortunately using the ISBN13 code wouldn't clear
    up the duplicate publisher issue, and might make things more
    complicated.

-   A quick dive to try to find a reason - could it be language,
    authors, date related? Look at a publisher with \>1 codes called
    "HarperCollins"

```{r}
books_isbn_pub %>% 
  filter(publisher == "HarperCollins") %>% 
  arrange(isbn13_publisher)
```

-   There is no obvious reason as to different publisher code used in
    the ISBN13 codes - I was hoping to see possibly a different language
    code or a pattern with number of authors or something similar.
