---
title: "W10 Day 2 Homework"
output: html_notebook
---

```{r message=FALSE}
library(tidyverse)
library(GGally)
```


# MVP

### 1. Load Data
```{r}
housing_prices <- read_csv("data/housing_prices.csv")

head(housing_prices)
```

<br>

### 2. Investigate total_rooms & total_bedrooms

```{r}
housing_prices %>% 
  select(total_rooms, total_bedrooms) %>% 
  ggpairs(, progress = FALSE)
```

It does look like there is a correlation between total rooms and total bedrooms

<br>


### 3. Drop total_bedrooms

```{r}
(housing_prices_trim <- housing_prices %>% 
  select(-total_bedrooms))
```


<br>

### 4 i. Investigate using `ggpairs`

```{r warning=FALSE}
housing_prices_trim %>% 
  ggpairs(progress = FALSE)
```

ii. Use `ggplot` on significant correlations with median_house_value

- I think there could be significant correlations with total_rooms, population, median income, ocean proximity


```{r}
housing_prices_trim %>% 
  ggplot(aes(x = median_house_value, y = total_rooms))+
  geom_boxplot()

housing_prices_trim %>% 
  ggplot(aes(x = median_house_value, y = population))+
  geom_boxplot()

housing_prices_trim %>% 
  ggplot(aes(x = median_house_value, y = median_income))+
  geom_boxplot()

housing_prices_trim %>% 
  ggplot(aes(x = median_house_value, y = ocean_proximity))+
  geom_boxplot()
```

<br>

### 5. Dummy variables

```{r}

housing_prices_trim %>% 
distinct(ocean_proximity)
```

There are 5 levels of ocean proximity, so we will need 4 dummy variables


<br>

### 6. SLR of median_house_value with median_income

median_house_value ~ median_income


```{r}
(model1 <- lm(median_house_value ~ median_income,
   data = housing_prices_trim))

# look at diagnostic plots before summary
autoplot(model1)
```
Plot 1: possible pattern of variation in residuals, but line is flat against 0 so would say good variation in residuals
Plot 2: QQ plot is a relatively straight line, therefore residuals are pretty normally distributed
Plot 3: a bit of evidence of heteroscedacity but not concerning 
**Not enough evidence to invalidate regression assumptions, this data works well for a linear regression model.**

```{r}
summary(model1)
```


Model results:
median_income is a significant term in our model given the small P value
median_income explains 41% of variance in median_house_value (see r squared is 0.4129)

RSE = 74870
Spread of data points is big

#### 7. Add another predictor

adding the ocean proximity variable

```{r}
(model2 <- lm(median_house_value ~ median_income + ocean_proximity,
   data = housing_prices_trim))

# look at diagnostic plots before summary
autoplot(model2)
```

Plot 1: not much pattern of variation in residuals, but line is flat against 0 so would say good variation in residuals
Plot 2: QQ plot is a relatively straight line, but diverts at end. Would say that residuals are pretty normally distributed
Plot 3: not much of evidence of heteroscedacity  


```{r}
summary(model2)
```

I would assume that house value would increase with proximity to the ocean
All levels of ocean proximity are significant
can see an improvement in r squared
