---
title: "Dplyr Homework KA"
output: html_notebook
---

```{r}
library(tidyverse)
```

```{r}
library(CodeClanData)
```

```{r}
drinks_content
```

# Question 1

#### Load in the dataset and have a look at it. What are the dimensions, variable types, variable names, etc.?

```{r}
dim(drinks_content)
```

-   The dimensions are 242 rows, 18 columns

```{r}
glimpse(drinks_content)
```

-   The variable types are character and double

```{r}
names(drinks_content)
```

-   The 18 variable names are: Beverage category, Beverage, Beverage prep, Calories, Total Fat (g), Trans Fat (g), Saturated Fat (g), Sodium (mg), Total Carbohydrates (g), Cholesterol (mg), Dietary Fibre (g), Sugars (g), Protein (g), Vitamin A (% DV), Vitamin C (% DV), Calcium (% DV), Iron (%DV), and Caffeine (mg)

# Question 2

#### Let's first investigate the calories of different drinks. Select the variables Beverage_category, Beverage, Beverage prep and Calories from your data. Since we are interested in the calorie content, check if there are any NA values in the data, and drop them if there are.

```{r}
drinks_content %>% 
  select(Beverage_category, Beverage, Beverage_prep,Calories) %>% 
  filter(is.na(Calories) !=TRUE)
```

-   There is no difference in the number of rows once `filter(is.na)` is applied so therefore there is no (NA) values in the Calories column to worry about!

# Question 3

#### Create a new variable (column) called calorie_diff, which stores the difference between 135 calories (135 calories = 10 cubes of sugar!) and the calories in each drink. (hint: you'll want to subtract 135 from the calories to see which drink have more or less than 10 cups of sugar).

```{r}
drinks_content %>%
  mutate(calorie_diff = Calories - 135, .after = Calories)


```

-   (Therefore there are 165 drinks that contain at least 10 cups of sugar) -

```{r}
drinks_content %>% 
  mutate(calorie_diff = Calories - 135, .after = Calories) %>% 
  filter(calorie_diff >= 0) %>% 
  select(Beverage, Calories)
```

# Question 4

#### Summarise the mean number of calories in each beverage_category. Which 3 category of drinks have the most calories? Which 3 drinks have the least? Write a small summary of your findings.

```{r}
drinks_content %>% 
  group_by(Beverage_category) %>% 
  summarise(Average_Calories = mean(Calories)) %>% 
  arrange(desc(Average_Calories)) %>% 
  head(3)
```

-   On average, the 3 most calorific drink categories are Smoothies, Frappuccino® Blended Coffees, and Signature Espresso Drinks

# Question 5

#### Let's look at this a different way. What is the average number of calories in each Beverage_prep type?

```{r}
drinks_content %>% 
  group_by(Beverage_prep) %>% 
  summarise(Average_Calories = mean(Calories))

```

# Question 6

#### Which Beverage_prep type contains more than the average calories of all drinks?

```{r}
  Average_Calories <- drinks_content %>% 
  summarise(mean(Calories)) %>% 
  pull() 
  drinks_content %>% 
  filter(Calories > Average_Calories) %>% 
    group_by(Beverage_prep) %>% 
  summarise()


```

-   There are 9 Beverage Prep types that contain more than average calories of all drinks.

# Question 7

#### Which is the best type of coffee drink (and by 'drink' we mean the combination of Beverage_category, Beverage and Beverage_prep) to get if you're worried about consuming too many calories?

```{r}
drinks_content %>% 
  filter(Beverage_category == "Coffee") %>% 
  select(Beverage_category, 
          Beverage,
          Beverage_prep,
          Calories) %>% 
  arrange(Calories) 
  
```

The coffee drink with the lowest calories is short, brewed coffee which contains 3 calories.
